---
title: "Homework #2 - deg2163"
author: "Diana Ballesteros Gonzalez"
date: "9/30/2018"
output: github_document
---

# **Problem 1**
```{r setup, include = FALSE}
## Set up 
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")
```


```{r load_library, message=FALSE}
## Loading the library
library(tidyverse)
library(ggridges)
```

```{r clean_data}
## Load the transit_data
transit_data = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", col_types = "cccddcccccccccccccccccccccccddcc")

## Clean the data 
transit_data_clean = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", col_types = "cccddcccccccccccccccccccccccddcc") %>% 
  janitor::clean_names() %>% 
  select(line:vending, ada) %>% 
  mutate(entry = tolower(entry),
    entry = recode(entry, "yes" = TRUE, "no" = FALSE))
```

### Step by Step Explanation
The variables contained in this new clean data file are line, station, name, station latitude / longitude, routes served (1 through 11), entry, vending, entrance type, and ADA compliance. The first thing I did was create a new dataset called "transit_data_clean" to distinguish it from the raw dataset. Then, I read in the raw dataset and input read_csv since it was a csv file using a relative path to make sure that I am able to pull in that data set if shared with others. Then, I used the janitor package to clean up the variable names to change the names to snake and lowercase. After, I selected only the variables that were requested by including all the columns from line to vending, then adding ADA compliance to the end. I then made all the rows within the entry variable lowercase by using tolower before recoding all the yes to equal 1 and no's to equal 0. The data is not yet tidy since some variables have not been recoded yet (e.g., vending).

### Problem 1 Answers
```{r transit_data_distinct_df}
## Created a new table with distinct stations 
transit_data_distinct = 
  distinct(transit_data_clean, station_name, line, .keep_all = TRUE) %>% 
  mutate(ada = as.logical(ada))
```
* The are `r nrow(transit_data_distinct)` distinct stations. 
* There are `r sum(transit_data_distinct$ada)` stations that are ADA compliant.
* The proportion of station entrances/exits without vending allow entrance is 
`r nrow(filter(transit_data_distinct, entry == TRUE, vending == "NO"))/nrow(filter(transit_data_distinct, vending == "NO"))`

```{r transit_data_reformat, message = FALSE}
## Reformatted data so that route number and route name are distinct variables. 
transit_data_distinct_reformat = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  distinct(transit_data_clean, station_name, line, .keep_all = TRUE) %>% 
  gather(key = "route_number", value = "route", route1:route11)
```

* There are `r nrow(filter(transit_data_distinct_reformat, route == "A"))` stations that serve the A train. Of the `r nrow(filter(transit_data_distinct_reformat, route == "A"))` stations that serve the A train, `r nrow(filter(transit_data_distinct_reformat, route == "A" & ada == TRUE))` are ADA compliant. 

# **Problem 2**
```{r import_datset}

## Importing Mr. Trash Wheels dataset
water_wheels_clean = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1, 
                     range = "A2:N336") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = round(as.integer(sports_balls)))
```

```{r precipitation_data}
## load the precipitation_2017 data 
precipitation_2017_clean = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, range = "A2:B14") %>% 
  janitor::clean_names()  %>% 
  mutate(year = 2017)

## load the precipitation_2016 data 
precipitation_2016_clean = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 5, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2016)

## binding the rows from both data sets 
precipitation_data = 
  bind_rows(precipitation_2016_clean, precipitation_2017_clean) %>% 
  mutate(month = month.name[month])
```

### Problem 2 Answers
* The number of observations in the water_wheels_clean data set is `r nrow(water_wheels_clean)`.
* The number of observations in the precipitation data set is `r nrow(precipitation_data)`.
* Some examples of key variables for Mr. Water Wheels are month, year, and weight_tons. Some examples of key variables for precipitation_data are month, year, and total precipitation during that year. 
* For available data, the total precipitation in 2017 was `r sum(precipitation_2017_clean$total)`
* The median number of sports balls in a dumpster in 2016 is `r water_wheels_clean %>% filter(year == "2016") %>% summarize(median(sports_balls))` 


# **Problem 3**
```{r setup_problem_3}
# install.packages("devtools")
devtools::install_github("p8105/p8105.datasets")

## loading the library and obtaining data
library(p8105.datasets)
data("brfss_smart2010")

## creating the brfss_smart2010_df
brfss_smart2010 = 
   brfss_smart2010 %>% 
   janitor::clean_names() %>% 
   filter(topic == "Overall Health") %>% 
   select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location)) %>% 
   spread(., key = response, value = data_value) %>%
   janitor::clean_names() %>% 
   mutate(excellent_very_good = excellent + very_good)
```

### Problem 3 Answers
* There are `r count(distinct(brfss_smart2010, locationdesc))` unique location included in the database
* Every state is represented. In fact `r count(distinct(brfss_smart2010, locationabbr))` states (including Washington DC) are represented. 
* The median of the “Excellent” response value is `r brfss_smart2010 %>% filter(year == 2002) %>% summarize(median(excellent, na.rm = TRUE))`

```{r problem_3_answers}
## A histogram of “Excellent” response values in the year 2002.
brfss_smart2010 %>% 
  filter(year == 2002) %>% 
  ggplot(aes(x = excellent)) + 
  geom_histogram(binwidth = 0.50, fill = "Blue") + 
  labs(
    title = "Excellent Response Values in 2002",
    x = "Excellent Data Values",
    y = "Frequency",
    caption = "Data from the brfss_smart2010 package"
  ) + 
  theme_bw() 

## Created a scatterplot
brfss_smart2010 %>% 
  filter(locationdesc == "NY - Queens County" | locationdesc ==  "NY - New York County") %>% 
  mutate(prop_excellent =  
    excellent / (excellent + fair + good + poor + very_good)
  ) %>% 
  ggplot(aes(x = year, y = prop_excellent, color = locationdesc)) +
  geom_point() + 
  labs(
    title = "Proportion of Excellent Responses from 2002-2010",
    x = "Year",
    y = "Proportion of Excellent Response",
    caption = "Data from the brfss_smart2010 package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE) + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

