---
title: "Homework #2 - deg2163"
author: "Diana Ballesteros Gonzalez"
date: "9/30/2018"
output: github_document
---

## Problem 1 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}

## do i have to add col_types for transit_data or transit_data_clean?

transit_data_clean = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:vending, ada) %>% 
  mutate(entry = tolower(entry),
    entry = recode(entry, "yes" = 1, "no" = 0))
```

### Explanation
The variables contained in this new clean data file are line, station, name, station latitude / longitude, routes served (1 through 11), entry, vending, entrance type, and ADA compliance. The first thing I did was create a new dataset called "transit_data_clean" to distinguish it from the raw dataset. Then, I read in the raw dataset and input read_csv since it was a csv file using a relative path to make sure that I am able to pull in that data set if shared with others. Then, I used the janitor package to clean up the variable names to change the names to snake and lowercase. After, I selected only the variables that were requested by including all the columns from line to vending, then adding ADA compliance to the end. I then made all the rows within the entry variable lowercase by using tolower before recoding all the yes to equal 1 and no's to equal 0. The data is not yet tidy since there are multiple variable 